### 2023-02-17 analysis/paper/paper.qmd
Edited by temporal pattern results & tried CV

---
title: "Variation in use of East Asian Late Paleolithic weapons: A study of tip cross-sectional area of stemmed points from Korea"
author:
  - Gayoung Park:
      institute: University of Washington
      correspondence: "yes"
      email: gayoungp@uw.edu
  - Marlize Lombard:
      institute: University of Johannesburg
  - Donghee Chong:
      institute: Kyung Hee University
  - Ben Marwick:
      institute: University of Washington
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
editor: visual
date: now
date-format: long
title-block-published: "Last updated"
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua

abstract: |
 The transition from the Early to Late Paleolithic in Korea is characterized by the introduction of blade technology, stemmed points, end scrapers, burins, denticulates, and finer grained materials. Stemmed points have been considered a representative tool that led this set of changes. In this study, we examine the possible role that stemmed points played during the technological transition, as well as throughout the Late Paleolithic period. Our main questions are: What were the best-fit ballistic probabilities for the stemmed points if they were hafted as weapon tips? How diverse were their likely uses? What are the temporal and spatial patterns of stemmed point use? We measured tip cross-sectional area (TCSA) to discriminate different likely use classes of projectile points, for example, as poisoned arrow tips or as stabbing spears. We analyzed TCSA with other variables including raw materials, weight, radiocarbon dates and locations. Our results show that the stemmed points could have served as javelin tips and stabbing spear tips, with smaller numbers as dart tips and un-poisoned arrow tips. TCSA values are controlled mostly by size rather than raw material types. We found different TCSA ranges of stemmed points at different sites, which could indicate people used stemmed points in different ways depending on the local environment. Some sites show a wide range of TCSA values that represent multi-purpose usage of stemmed points. The temporal pattern of TCSA values is one of little change throughout the Late Paleolithic period but points were predominantly produced before the Last Glacial Maximum (LGM). We observed that stemmed points were mostly located in certain ecoregions in Korea but no clear spatial pattern was confirmed. We conclude that stemmed points were multi-functional tools, with many likely designed for javelin and stabbing spear tips.

keywords: |
  Stemmed points; Korean Paleolithic; Tip Cross-Sectional Area
---

# Introduction {.unnumbered}

The introduction of new stone artifact technologies marks a major transition in the Korean Paleolithic from the Early to the Late Paleolithic periods. The transition includes blade technology, stemmed points, end scrapers, burins, denticulates, etc. [@bae2017origin; @nakazawa2018quaternary; @bae2017late; @lee2017korean; @seong2016eastern]. Stemmed points are considered to be the first evidence of a suite of new technologies defining the Late Paleolithic period in this region [@seong2016eastern; @seong2008tanged]. This is related to the fact that stemmed points appear to originate from Korea and spread throughout Northeast Asia, and they have a close association with mobility, site formation, and occupation diversity [@o2018origins; @Chong_2021; @park2022did]. Despite the importance of stemmed points, only a few studies to date have examined their likely uses. Previous work has mostly discussed their origin, the chronology of the Korean Late Paleolithic, and their relationship with the Japanese archipelago [@chang2013human; @Park_2013; @lee2019were; @Chong_2021].

The purpose of this study is to examine the possible uses of stemmed points in order to understand how stemmed points played a role in the technological transition. We use the tip cross-sectional area (TCSA) metric to infer potential variability in weapon-use strategies, and compare the results to other archaeological and ethnographic assemblages based on the idea that different shapes correspond to different weapon types [@lombard2021variation]. We then explore the relationship between the TCSA range and raw materials, artifact size and discard location, and how these change over time. Our main questions are: What were the best-fit ballistic probabilities for the stemmed points if they were hafted as weapon tips? How diverse were their likely uses? What are the temporal and spatial patterns of stemmed point uses? As part of our discussion, we examine possible links between the roles of stemmed points and environmental change or population dynamics during the Late Paleolithic period. In order to understand how a certain weapon-tip type was chosen, we apply an evolutionary perspective with the assumption that people chose their weapon tip types as part of their adaptation to the socio-environmental circumstances they found themselves in.

## Stemmed Points in Korea and East Asia

Stemmed points (Sumbejjirugae in Korean) are projectile points made on an elongated blade-like flake or blade with two parallel facets and a single or two arises that converge to form an inverted "Y" [@pratt2020circum] (@fig-sp-hand-made). Slight retouch is typically performed on the proximal end to shape an acute tip and on the distal end to make a stem, which connects to a wooden shaft. Elsewhere in the world these types of artifacts are often called 'tanged points', but we prefer referring to them as 'stemmed points' here to distinguish them from Bronze Age stone projectile points which have long been called 'tanged points' in Korea [@park2022did]. Understanding the appearance of stemmed points is relevant to general questions about the origins of projectile technology, the technological transition of the Late Paleolithic, and relationships between Korea and adjacent regions in East Asia, such as Japan.

Stemmed points are the first composite tools in the Korean Paleolithic that required two different parts to be one complete tool, a stone point and a shaft, presumably made out of wood [@seong2008tanged]. Using blades as the blank for the point enables mass production of this composite tool and its shape can become more standardized [@lee2015characteristics; @park2022did]. Therefore, as @o2018origins claimed, understanding the emergence of projectile technology provides insights into greater cultural, evolutionary, and behavioral cognitive flexibility.

Since stemmed points represent the beginning of the Korean Late Paleolithic, investigating their origins is critical for understanding the technological transition from the Early Paleolithic, modern human dispersals into the region, and claims for the existence of the 'Middle' Paleolithic in Korea [@bae2010origin; @bae2017late; @norton2009evolution; @seong2016eastern]. Debate about explaining the origin of stemmed points which can be summarized into two competing models: a 'heterogenic' migration [@bae2010origin], and an *in situ* evolution model [@seong2009emergence]. The migration model claims that the new blade industry including stemmed points, and the earlier coarse flake tradition including large cores, polyhedrons, choppers, and even handaxes, came from different origins (i.e. 1: North route: Siberia, Mongolia, or other regions of northeast China, and 2: South route: southern China) as the result of a continuing influx of modern human migration from two routes. The *in situ* model argues that stemmed points and other Late Paleolithic technologies, such as blade industries, autonomously emerged in the south of the Korean peninsula, with no apparent external influence. The difference between the two models comes from how they explain a few early sites that contain stemmed points but no blades and blade cores. While the *in situ* model claims that stemmed points appeared before blades by acknowledging those early sites, including Bonggok, Songamri, Yonghodong, and Hwadaeri, the migration model supports the traditional definition of the Late or Upper Paleolithic in Europe and Africa that started with the emergence and establishment of blade technology [@seong2016eastern].

Stemmed points are important proxies to understand human behaviors during the Late Paleolithic. @park2022did examined mobility and site occupation patterns by applying concepts of human behavioral ecology to lithic assemblages. The study found that forager groups using stemmed points may have been associated with the occupation of marginal or extreme environments, compared to groups with no stemmed points. Also, stemmed points were more frequently associated with expedient technologies, indicating residential and less mobile behaviors. @Chong_2021 claims that the morphological variation of the stemmed points along with tool types in lithic assemblages, assemblage size, use of raw materials, and types of blanks could represent specific characteristics of occupation such as a "limited activity station" and a "residential base camp." For example, stemmed points with high morphological variations in tool size, shape of edge, degrees of damage, and types of edge retouching from Yongsandong site may indicate that the site was used for specific or limited activities such as hunting [@kim2004yongsan; @seong2015diversity; @Chong_2021].

The connection between stemmed points in both Korea and Japan has been studied since the late 1980s as a part of evidence for long-distance/maritime cultural interchanges or social networks [@lee2015characteristics; @chang2013human]. Stemmed points from the Bonggok site in Korea are currently accepted as the oldest ones among Northeast Asia, dated to ca. 41.5 ka, and made out of elongated flakes [@seong2009emergence; @seong2015diversity; @bae2017origin]. After their first appearance in Korea, stemmed points (Hakuhensenntouki in Japanese) appeared in Kyushu, Japan during the late Marine isotope stage (MIS) 3. In addition to the stemmed points, there are similar artifacts found in both regions such as microblade cores, Moppule-seokgi (Kakusuijyosekki  in Japanese), backed knives, bilateral points, bifacial points, and transport of obsidian [@lee2012characteristics; @chang2013human; @lee2015characteristics; @kim2021evidence].

```{r}
#| label: fig-sp-hand-made
#| fig-cap: "Stemmed points from Yongsandong site. 1-6: plain stemmed points, 7-8: one side denticulate stemmed points."

knitr::include_graphics(here::here("analysis/figures/hand-made-figures/TCSA_sp.png"))
```

## Previous studies about the function of stemmed points

Though stemmed points have generally been assumed to have been hunting armatures [@lee2019were; @lee2011study; @seong2008tanged; @lee2002new; @chang2013human], it is difficult to determine their likely uses without knowing their complete shape when attached to other components. Preserved wooden components of prehistoric projectile tools are too few and rare to standardize their overall scale and variability [@shea2006origins]. @Lee_Kong_2002 claimed stemmed points should be considered more generally as 'stemmed tools' because of their diverse shapes, which could represent different functions, and the uncertainty of their complete shape. Other researchers propose that non-symmetrical stemmed points, with retouch on one side or denticulate blades on one side, should not be referred to as stemmed points but rather as stemmed knives, stemmed side-scrapers, stemmed end-scrapers, or stemmed burins (@fig-sp-hand-made: 7-8) [@seong2008tanged; @Lee_Jang_2011; @Kim].

Stemmed points are typically symmetrical from tip to tang, with the central axis serving as a line of symmetry [@Lee_Jang_2011]. There is a high percentage of broken tips and stems, and the reused tools were repaired in accordance with symmetry [@Park_2013; @Kim_2017]. Studies of stemmed point manufacturing processes and the patterns of broken pieces show that stemmed points may have been used mainly as spear tips [@Lee_1985; @Chang_2002]. For example, at Yongsandong site, only 10% of the tools are complete, while 33% of the tips are missing. In the case of Jingeuneul, the percentages are 16% and 50%, respectively [Park_2013]. In addition to the morphological aspect of stemmed points, investigations of a whole site and the tool composition of an assemblage suggest that stemmed points or stemmed tools could be strongly associated with hunting activities including peeling the animal skin after slaughtering or separating the bones from the flesh [@seong2008tanged; @Chong_2021].

## Tip Cross-Sectional Area

Tip cross-sectional area (TCSA) of stone artifacts has been used as a ballistically relevant standard to probabilistically discriminate different likely use classes of projectile armatures such as spearthrower (a.k.a. atlatl) dart tips, un-poisoned arrow tips and large stabbing/thrusting spears [@sisk2011african; @lombard2021variation; @lombard2020tip; @lombard2021did; @o2018origins; @lombard2022standardized]. It is critical to note that the TCSA metric alone cannot unambiguously determine artifact function; it only suggests a best-fit ballistic probability for the points if they were hafted as weapon tips. The TCSA metric represents the part of the tool that cuts the target's hide and is related to weapon flight and penetration dynamics [@hughes1998getting; @sitton2020tip; @lombard2021variation]. This method was first proposed by @hughes1998getting and validated by @shea2006origins through the comparison of archaeological examples with ethnographically collected samples of known use. @shea2009impact also applied this approach to compare projectile points from Africa, the Levant, and Europe, and claimed that projectile weapons first appeared in Africa.

One of the key advantages of the TCSA metric is its convenience of application: regardless of the point type, only the maximum width and thickness measurements are required to calculate the TCSA value (0.5 x maximum width x maximum thickness) [@sisk2011african; @lombard2020tip]. Later, @sisk2011african proposed an alternate metric, tip cross-sectional perimeter (TCSP), for a more precise measure of the force required to penetrate a target to a lethal depth, whereas the TCSA metric is more associated with cutting. However, TCSP has a few disadvantages that limit its applicability to our case study of Korean stemmed points. The force and penetration depth are not only affected by the stone tip, but also by the mass of the shaft, which cannot be known for most archaeological stone-tipped weapons because they have not been preserved in the archaeological record [@lombard2020tip]. @sisk2011african also mentioned that TCSP cannot be applied to backed pieces that were hafted as projectile armatures.

@lee2019were first applied TCSA to stemmed points from Korea along with use-wear analysis. They analyzed stemmed points from Jingeuneul, located in the southwest of Korea, which has the largest number of stemmed points (n = 99) found at a single site to date. For the TCSA, they were only able to use ten stemmed points since they selected stemmed points that retained the widest and thickest parts of the specimens and showed diagnostic impact fractures for the use-wear analysis. The purpose of using TCSA in their research was to compare the values to North American dart tips and arrowheads. Their results show that the range of TCSA values for Jingeuneul stemmed points is relatively wide, overlapping with both North American dart tips and arrowheads. According to their use-wear analysis, a significant number of the stemmed points have diagnostic impact fractures (DIFs) on the surface, likely caused by longitudinal forces from the shaft. Based on the results, they conclude that stemmed points may have been used as spear-throwers or bows. Inspired by @lee2019were and TCSA research from lithic assemblages in other parts of the world, we aim to investigate TCSA values for a much larger sample of stemmed points from all over South Korea to better understand their likely uses during the Late Paleolithic.

# Methods

## Archaeological sites

```{r}
#| label: import-data-and-compute-tcsa

library(readxl)
library(here)
library(tidyverse)
library(dplyr)
library(ggbeeswarm)
library(ggplot2)

# import raw data
mydata <- read_excel(here("analysis/data/raw_data/TCSA_raw_data.xlsx"))

# compute TCSA
tcsa <- mydata %>%
  mutate(min_tcsa = 0.5 *Width * Thickness) %>%
 # mutate(max_tcsa = 0.5 *Length * Thickness) %>%
  dplyr::filter(min_tcsa <= 250) %>% #omit very large TCSA, could be measurement error
  # group rare raw materials < 10
  mutate(raw_materials_top = fct_lump_min(Raw_material, 10)) %>%
  mutate(site_name_top = fct_lump_min(Site_name, 5))

```

```{r}
#| label: fig-map
#| fig-cap: "Korean Paleolithic sites mentioned in this study"
#| fig-width:   4.45
#| fig-height:  4.5

# Facet by raw materials

korean_archaeological_site_locations <-
  read_excel(here::here("analysis/data/raw_data/TCSA_site_locations.xlsx"))

korean_archaeological_site_locations <-
  tibble::rowid_to_column(korean_archaeological_site_locations, "ID")

library(ggmap)
library(ggrepel)
# devtools::install_github('3wen/legendMap')
library(legendMap)

# if we have already downloaded the map tiles, skip doing it again
if (file.exists(here::here("analysis/data/raw_data/map.RData"))) {

  # load the the map tiles have already been downloaded
  load(file = here::here("analysis/data/raw_data/map.RData"))

  } else {

    # download map tiles for our study area if we don't 
    # have them already
map <-
  get_stamenmap(bbox = c(left = 124,
                         bottom = 33,
                         right = 	130,
                         top = 39),
                zoom = 10)
# and save the tiles so we don't have to do this every time we render
save(map, file = here::here("analysis/data/raw_data/map.RData"))

}

# map with site names next to the points
map_of_sites <-
  ggmap(map)  +
  geom_point(data = korean_archaeological_site_locations,
             aes(long_dd ,
                 lat_dd),
             colour = "red",
             size = 2) +
  # label points with site name
  geom_text_repel(data = korean_archaeological_site_locations,
                  aes(long_dd ,
                      lat_dd,
                      label = Site_name),
                  size = 2,
                  bg.color = "white",
                  bg.r = 0.15) +
  theme_minimal(base_size = 6) +
  labs(x = "Longitute",
       y = "Latitude") +
  legendMap::scale_bar(
    # edit these numbers to select a suitable location
    # for the scale bar where it does not cover
    # important details on the map
    lon = 124.5,
    lat = 33.5,
    legend_size = 2,
    # distance of one section of scale bar, in km
    distance_lon = 50,
    # height of the scale bar, in km
    distance_lat = 5,
    # distance between scale bar and units, in km
    distance_legend = 20,
    # units of scale bar
    dist_unit = "km",
    # add the north arrow
    orientation = TRUE,
    # length of N arrow, in km
    arrow_length = 60,
    # distance between scale bar & base of N arrow, in km
    arrow_distance = 50,
    # size of letter 'N' on N arrow, in km
    arrow_north_size = 10) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

# inset world map, from
# https://stackoverflow.com/a/62104396/1036500
library(maps)
library(magrittr)
library(maptools)
library(raster)
library(ggplot2)

#Defining a general CRS
mycrs <- "+proj=longlat +datum=WGS84 +no_defs"

#Using the original maps package, then converting map into SpatialPolygons object
world <- maps::map("world", fill=TRUE,  plot = FALSE) %$%
  maptools::map2SpatialPolygons(., IDs=names,proj4string=CRS(mycrs))

#The resulting map has self intersection problems so any further operation reports errors; using buffers of width 0 is a fast fix
while(rgeos::gIsValid(world)==FALSE){
  world <- rgeos::gBuffer(world, byid = TRUE, width = 0, quadsegs = 5, capStyle = "ROUND")
}

#Dissolving polygon's limits
world <- raster::aggregate(world)

worldplot <-
  ggplot() +
  geom_polygon(data = world,
               aes(x=long, y=lat,
                   group=group),
               fill='NA',
               color='black',
               size=0.1) +
  annotate("rect",
           xmin = 118,
           xmax = 137,
           ymin = 28,
           ymax = 47,
           fill = NA,
           colour = "red",
           size = 0.5
  ) +
  theme_void() +
  coord_fixed(1) +
  labs(x = "", y = "") +
  labs(x = NULL, y = NULL, title = NULL) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(panel.background = element_rect(fill = "white"),
        panel.border = element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

# combine site map and inset map
library(cowplot)
ggdraw() +
  draw_plot(map_of_sites) +
  draw_plot(plot = worldplot,
            x = 0.45, # x location of inset placement
            y = -0.27, # y location of inset placement
            width = .6, # Inset width
            height = .75, # Inset height
            scale = 0.5 # Inset scale
  ) +
  theme(panel.background = element_rect(fill='white', colour="white"),
        plot.background = element_rect(fill='white', colour="white"))

```

```{r}
# move dataframes from the Results section to create codes for numbers in the text

tcsa_site <- tcsa %>% 
  group_by(site_name_top) %>% 
  add_tally() %>% 
  filter(n>1) %>% 
  mutate(label_site = as.factor(paste0(site_name_top,' (n = ', n, ")"))) %>%   
  dplyr::select(min_tcsa,
         site_name_top, 
         label_site,
         n) 
# %>% 
#   distinct_all()

# get rid of 'Other'
tcsa_site_num <- tcsa_site %>% 
  filter(site_name_top != 'Other')

###########
# radiocarbon data per assemblage
sitedate <- read_excel(here("analysis/data/raw_data/TCSA_site_info.xlsx"))

mydata_ages <-
  sitedate %>%
  separate(Period, into = c('age', 'error'),
           sep = "±") %>%
  mutate(age_ka = parse_number(age) / 1000,
         error = parse_number(error))

# combine the assemblage with age data

tcsa_site_ass <- tcsa %>% 
  left_join(mydata_ages) %>% 
  group_by(Assemblage) %>% 
 drop_na(age_ka) %>% 
  add_tally() %>% 
  mutate(label_site = as.factor(paste0(Assemblage,' (n = ', n, '), (', {round(age_ka,1)}, 'ka)' ))) %>%   
  dplyr::select(age_ka,
         min_tcsa,
         Assemblage, 
         label_site,
         n) %>% 
    distinct_all()

# raw material : create column for "n = xx "
tcsa_raw <- tcsa %>% 
  group_by(raw_materials_top) %>% 
  add_tally() %>% 
  filter(n>1) %>% 
  mutate(label_raw = as.factor(paste0(raw_materials_top,' (n = ', n, ")"))) %>%   
  dplyr::select(min_tcsa,
         raw_materials_top, 
       label_raw,
       n) %>% 
  distinct_all()

# remove NAs for weight
tcsa_mass_no_na <-
  tcsa %>%
 drop_na(`Weight(g)`)

# drop na for dates
mydata_ages_na <- mydata_ages %>%
  drop_na(age)

# excluded artifact #
tcsa_omit <- mydata %>%
  mutate(min_tcsa = 0.5 *Width * Thickness) %>%
  dplyr::filter (min_tcsa > 250)

# code for numbers in the text
total_site <- nrow(korean_archaeological_site_locations)
total_sp <- nrow(mydata)
total_ass <- nrow(mydata_ages)
n_sites <- length(unique(tcsa$Site_name))
min_n_points <- nrow(tcsa_site) 
# n_radiocarbon_ages <- NULL # we only have one age per assemblage
n_sites_with_radiocarbon_ages <- nrow(mydata_ages_na)
n_omit <- nrow(tcsa_omit)
n_ass <- length(unique(tcsa$Assemblage))
n_raw <- nrow(tcsa_raw)
n_size <- nrow(tcsa_mass_no_na)
# n_spatial <-nrow(tcsa_site_num) # got rid of "Other" in the dataframe
# n_spatial_ass <- length(unique(tcsa_site_num$Assemblage))
```

After the first discovery at the Seokjangri site in the 1960s, more than 450 stemmed points have been found at over 30 sites throughout South Korea [@Sohn_1967; @lee2019were; @Chong_2021]. Most sites contain only a few stemmed points and only a few sites have many more, such as Jingeuneul, Suyanggae (n = 55), and Yongsandong (n = 38) [@Kim_2017]. Among these stemmed points, we selected those that retained their widest and thickest parts. We included stemmed points discovered during field surveys as well as those found at sites that were never dated but were associated with other Late Palaeolithic artifacts. Applying these sampling criteria resulted in a sample of `r total_sp` stemmed points from `r total_ass` assemblages unearthed from `r total_site` sites spanning the period 44-10 ka (@fig-map). The dimensions of the `r total_sp` stemmed points were obtained from published excavation reports and by direct measurements during our visits to the collections of local museums and archaeological institutions in Korea.

We distinguished between multiple assemblages at a single site where numerous excavations have taken place in different locations at the site, and by different institutions under the same site name. For example, Suyanggae site, a registered Korean National Heritage site, has been excavated more than ten times since 1980 by the local university museum and later by archaeological research institutions. There are six different excavation locations that range from a few meters apart to a few kilometers. Similarly, we identified multiple assemblages in a single excavation or even a trench where archaeological deposits were separated by culturally sterile deposits, or where distinct artifact-bearing stratigraphic units could be identified by major differences in the texture, color, and composition of the sedimentary deposits. As a result, in this research we used four assemblages from Suyanggae. We separated one assemblage from the four by using a different site name, Hajinri, following the convention established by the excavators at that location. Hajinri is the sixth excavation location at Suyanggae, which is 3.5 kilometers apart from the other areas and dated much earlier (around 42-30 ka) than the other assemblages (around 31-15 ka). According to the excavation reports for Hajinri, stemmed points first appeared there around 42 ka with the earliest blades and blade cores [@Lee_Woo_Lee_An_Yun_Park_Otani_Kim_Kim_Han_et_al_2018]. While we include data from Hajinri here, we consider this assemblage as an unusual outlier because the stemmed points and blades from Hajinri are highly standardized and refined, which are only found at other sites much later in time. We have more confidence in finds from Bonggok, which has the second-earliest dates (around 41.5 ka) in our collection, as the first appearance of stemmed points. Bonggok includes blades or elongate flakes, but without any accompanying blade cores [@park2022did].

## Investigating patterns in TCSA values

To answer our research questions about the likely uses for stemmed points, we calculated TCSA values for the stemmed points in our sample. We also explored the interaction of TCSA values with raw materials and artifact size, which we used weight as a proxy. Because the shape of stone artifacts is highly influenced by the raw materials used [@mcpherron2014experimental]. It is also assumed that raw materials are highly correlated with tool size. We investigated radiocarbon ages associated with the points in order to determine temporal patterns in TCSA values. We separated the research time period into three phases based on the major climate event during the Late Pleistocene, the Last Glacial Maximum (LGM), and examined the distribution of TCSA values before, during, and after the LGM. We explored the relationship between TCSA values and the location of assemblages by comparing the distributions of TCSA values across ecological and vegetation zones.

```{r}
#| label: tbl-tcsa-ranges
#| tbl-cap: "TCSA range from Lombard (2021 & 2022)"

library(gt)
library(sass)

read_csv(here::here("analysis/data/raw_data/tcsa-range-from-lombard-2022.csv")) %>% 
  knitr::kable(caption = "TCSA range from Lombard (2021 & 2022)")
```

To aid with interpreting our results, we followed the TCSA ranges for different weapon-delivery systems that @lombard2022tip and @lombard2021variation created by summarizing the analysis by @wadley2008segment of backed microliths (@tbl-tcsa-ranges). We excluded `r n_omit` artifacts from our dataset with TCSA greater than 250 and thus outside of the range of the comparative data. For this comparison with weapon-delivery systems we include TCSA values from a total of `r min_n_points` stemmed points from `r n_ass` assemblages unearthed from `r n_sites` sites.

### Raw materials

Selective use of raw materials is a key characteristic of the technological transition of the Korean Late Paleolithic. Prior to the Late Paleolithic, people mostly used quartzite and vein quartz for stone artifacts. Then, finer grained materials were added to assemblages for producing the newly introduced tools [@seong2004quartzite]. We analyzed TCSA values of `r n_raw` artifacts with raw material information to examine the interaction of raw material types and TCSA values. We categorized rare raw materials, which have less than 10 artifacts, as "Other".

### Weight

Different sizes of stone artifacts can constrain or enable different functions. Overall size of stone points has been used as a potential proxy for identifying different types of armatures [@thomas1978arrowheads; @sahle2019assessment]. We chose weight as a proxy for the overall size of the stemmed points. As a reliable discriminator between tools of different sizes as well as a descriptive attribute, weight can be measured rapidly and objectively [@fenenga1953weights; @shea2006origins]. We then explored the relationship between weight and raw materials on TCSA values. Excluding points that we were not able to directly measure or obtain records of their weight, we explored the relationship between TCSA and weight for `r n_size` artifacts.

### Temporal patterns

We used radiocarbon ages to investigate the temporal patterns of the likely uses of stemmed points. After excluding assemblages that have no radiocarbon dates, we used `r n_sites_with_radiocarbon_ages` assemblages dated from 45ka to 14.8ka to explore variation in TCSA over time. We divided the artifacts from dated assemblages into three groups based on the LGM: before, during and after, to examine the impact of this major climate event on TCSA values.

### Spatial Patterns

We summarized the distribution of TCSA values for each assemblage to analyze spatial patterns in TCSA values. Among `r n_sites` sites, we combined those that contain fewer than five stemmed points and named them "Other". We explored the possible effect of environmental variation on TCSA values by grouping stemmed points by the eco-regional zones that they were found in, and comparing the distributions of TCSA values across the different zones. We compared the distributions across four vegetation zones and 16 zones that differed in terms of their geographical characteristics.

The vegetation zones are based on @yi2010vegetation's classification of South Korea into three zones: Central Temperate Zone (CT), South Temperate Zone (ST), and Subtropical-warm Temperate Zone (SWT). These divisions are based on @yim1975distribution's forest vegetation map, defined by recent temperature and precipitation values.

We also explored spatial patterns using 16 geographical zones based on geographical characteristics including inland, coastal areas, major rivers, islands, and major mountain ranges in addition to ecological information including temperature and precipitation [@lee2008relationship]. The sites in our sample occur in 14 of these zones: Imjin river basin (IRB), Metropolitan (MP), Central inland (CI), Kangwon coastal (KC), Choongnam coastal (CC), Southwestern inland (SWI), Upper Nagdong river basin (UNRB), WoolYoung coastal (WYC), Western Cholla (WC), Southern mountain (SM), Southeastern inland (SEI), Hyungsan Taewha coastal (HTC), Western south coastal (WSC), and Eastern south coastal (ESC). Then, we used longitudes and latitudes to locate individual assemblages and examined TCSA values for each zone. We fully acknowledge that modern eco-regional maps may not fully resemble the Pleistocene landscape in which the points were made and used, however these are the only maps of eco-regional zones that are currently available.

## Modelling weapon-tip type selection

The process of introducing novel technologies can vary based on the social and environmental context in which the transmission of manufacturing techniques for the new tools takes place. Furthermore, the new tools themselves can alter the typical contexts of tool use. In America, for example, the advent of bows and arrows provided non-elite hunters with the opportunity to produce their own subsistence or pursue individual wealth without the necessity of hunting in teams [@bettinger2013effects; @angelbeck2014faustian; @rorabaugh2015timing]. Taking an evolutionary approach, we assume that, given an opportunity to explore alternative technologies, human groups carefully selected a specific stone tool technology based on its advantages over other alternatives, according to their performance in a variety of domains, such as physical and social functions [@lombard2022tip]. Thus, the selection of weapon-tip types is likely to reflect the socio-environmental circumstances that people encountered and managed. In one example, @eren2022plains compared the morphological variance of Clovis and Folsom points and claimed that Clovis points were more variable in shape than Folsom points because Clovis foragers were exposed to largely unfamiliar landscapes. Clovis points were used as multifunctional tools that performed a wider range of tasks, including cutting and sawing. Folsom points, on the other hand, show a narrower range of variation, indicating they were more likely used for specific tasks.

Inspired by @eren2022plains's approach, we hypothesized some simple scenarios that might explain the temporal and spatial patterns of TCSA range in Korea. If stemmed points have a narrow range of TCSA values, then people likely produced tools that performed a small set of specific tasks. This may be related to low levels of uncertainty in the forager's social and physical environments. On the other hand, a wide range of TCSA values may indicate that stemmed points were multifunctional tools, suggesting that people were responding to unfamiliar situations, such as moving into an unfamiliar landscape or unpredictable variation in patch productivity and travel times [@bird2006behavioral; @bettinger2016marginal; @kelly2007foraging].

We predict a temporal pattern of more variable TCSA values at the first appearance of stemmed points, suggesting that the tools were being used as part of an adaptation to moving into unfamiliar landscapes, with unpredictable variations in patch productivity and travel times. The variability of TCSA values in the LGM is predicted to increase as lower temperatures alter the distribution of resources and reduce the predictability of resource encounters. After the LGM we anticipate a reduction in the range of TCSA values, with higher temperatures and increased bioproductivity.

In response to heterogeneously distributed food resources, TCSA values are predicted to vary across vegetation types and geographical zones. In harsh environments with lower patch productivity, similar to the case of LGM duration, we expect to observe a further increase in the variability of TCSA values. On the other hand, we predict a reduction in the range of TCSA values in affluent patches with predictable types of prey. In their study of hunter-gatherer mobility strategies, @hamilton2016ecological show that resources are abundant and predictable along coasts or lake shores, which is highly dependent on temperature and precipitation and thus hunter-gatherers often become effectively sedentary. Therefore, our prediction is that the range of TCSA values will be narrower in coastal areas such as vegetation zones ST and SW and geographical zone ESC, compared to inland areas such as vegetation zone CT and geographical zone CI.

The entire R code [@core2021team] and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/10.17605/osf.io/dqna8 to enable re-use of materials and improve reproducibility and transparency [@marwick2017computational]. All of the figures, tables, statistical test results presented here can be independently reproduced with the code and data in this repository. The code is released under the MIT license, the data as CC-0, and the figures as CC-BY, to enable maximum re-use.

# Results

## TCSA range of Korean stemmed points

@fig-tcsa-all-sp shows TCSA values for all stemmed points in our sample, with shaded rectangles to assist in the interpretation of their likely uses. Overall we see a wide variation in TCSA values. The sample mean of TCSA is `r round(mean(tcsa$min_tcsa), digits=1)`, and the standard deviation of TCSA is `r round(sd(tcsa$min_tcsa), digits=1)`. According to the TCSA ranges presented in @tbl-tcsa-ranges, Korean stemmed points are mostly found in the categories of javelin tips and stabbing spear tips, with smaller numbers as dart tips and un-poisoned arrow tips. Among the weapon-tip types, only poison arrow tips appear to be absent from these Korean assemblages.

```{r}
#| label: fig-tcsa-all-sp
#| fig-cap: "Distribution of TCSA values for all Korean stemmed points in the current dataset. Shaded boxes indicate TSCA ranges for various weapon types."
#| fig-width:   3
#| fig-height:  2

# all artifacts overall

point_function_group_label_y <- 0.75
point_function_group_label_size <- 1.5

ggplot(tcsa,
        aes(0.3, 
            min_tcsa)) +
  # show regions for tools from Lomabrd 2021 & 2022
  # https://www-sciencedirect-com.offcampus.lib.washington.edu/science/article/pii/S0277379121002286
  # https://assets.researchsquare.com/files/rs-2016964/v1/00d381a3-3ecf-4d5b-b2ec-89a41227bbd5.pdf?c=1662475267
    annotate("rect", 
           xmin = 0.1,
           xmax = 0.2,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Poisoned arrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (18-4) /2 + 4) +
      annotate("rect", 
           xmin = 0.2,
           xmax = 0.3,
           ymin = 17,
           ymax = 47,
           fill = "blue",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Un-poisoned arrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (47-17) /2 + 17) +
      annotate("rect", 
           xmin = 0.3,
           xmax = 0.4,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Dart tips",
               hjust = 0,
           x = point_function_group_label_y,
           y = (76-40) /2 + 40 - 5) +
      annotate("rect", 
           xmin = 0.4,
           xmax = 0.5,
           ymin = 42,
           ymax = 90,
           fill = "purple",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Javelin tips",
               hjust = 0,
           x = point_function_group_label_y,
           y = (90-42) /2 + 42 ) +
     annotate("rect", 
           xmin = 0.5,
           xmax = 0.6,
           ymin = 80,
           ymax = 200,
           fill = "yellow",
           alpha = 0.2) +
      annotate("text", 
               label = "Stabbing spear tips",
               size = point_function_group_label_size,
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (200-80) /2 + 80 ) +
  geom_boxplot(alpha = 0.1) +
  geom_quasirandom(alpha = 0.4,
                   size = 0.7) +
  coord_cartesian(xlim = c(-0.1,1)) +
  theme_bw(base_size = 6) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()
        ) +
  labs(y = "TCSA")
```

## Variation in TCSA values by artifact size

```{r}

# Facet by size class
# https://cran.r-project.org/web/packages/Ckmeans.1d.dp/vignettes/Ckmeans.1d.dp.html
library(Ckmeans.1d.dp)

# find clusters
result <- Ckmeans.1d.dp(tcsa_mass_no_na$`Weight(g)`)

tcsa_mass_no_na <-
tcsa_mass_no_na %>%
  mutate(mass_cluster = factor(result$cluster))

```

Using weight as a size proxy we examined the relationship between size and likely use of the stemmed points inferred from TCSA values. We conducted a univariate cluster analysis [@song2020efficient; @wang2011ckmeans] of stemmed points by weight, revealing three clusters of artifact sizes (A of @fig-artifact-size-cluster, mean = `r round(mean(tcsa_mass_no_na$'Weight(g)'), digits=1)`, SD = `r round(sd(tcsa_mass_no_na$'Weight(g)'), digits=1)`). Cluster 1, the smallest (lightest) artifacts, has a lower range of TCSA values, compared to Cluster 2 (B of @fig-artifact-size-cluster, mean = `r round(mean(tcsa_mass_no_na$'min_tcsa'), digits=1)`, SD = `r round(sd(tcsa_mass_no_na$'min_tcsa'), digits=1)`). TCSA values for Cluster 3 are the highest, except for one artifact, which is lower than 50. There are a few overlaps, but we can still assume that artifacts in different size clusters might have been made for different purposes.

```{r}
#| label: fig-artifact-size-cluster
#| fig-cap: "A. Distribution of artifact weight showing three clusters. B. TCSA values for all artifacts. Artifact size classes indicated by the digits representing data point values"
#| fig-width:   4
#| fig-height:  2

plot_mass_cluster <- 
ggplot (tcsa_mass_no_na,
        aes(1, 
            `Weight(g)`)) +
  geom_boxplot(outlier.shape = NA) +
  geom_quasirandom(aes(shape = mass_cluster),
                   alpha = 0.7,
                   size = 1.1) +
  scale_shape_identity() +
  labs(shape='Cluster',
       y = "Weight (g)") +
  theme_bw(base_size = 6) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()
        )

plot_tcsa_mass_cluster <- 
ggplot (tcsa_mass_no_na,
        aes(0.3, min_tcsa)) +
  geom_boxplot(outlier.shape = NA) +
  geom_quasirandom(aes(shape = mass_cluster),
                   alpha = 0.7,
                   size = 1.1) +
  scale_shape_identity() +
  labs(shape='Cluster') +
  theme_bw(base_size = 6) +
    annotate("rect", 
           xmin = 0.1,
           xmax = 0.2,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Poisoned arrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (18-4) /2 + 4) +
      annotate("rect", 
           xmin = 0.2,
           xmax = 0.3,
           ymin = 17,
           ymax = 47,
           fill = "blue",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Un-poisoned arrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (47-17) /2 + 17) +
      annotate("rect", 
           xmin = 0.3,
           xmax = 0.4,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Dart tips",
               hjust = 0,
           x = point_function_group_label_y,
           y = (76-40) /2 + 40 - 5) +
      annotate("rect", 
           xmin = 0.4,
           xmax = 0.5,
           ymin = 42,
           ymax = 90,
           fill = "purple",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Javelin tips",
               hjust = 0,
           x = point_function_group_label_y,
           y = (90-42) /2 + 42) +
     annotate("rect", 
           xmin = 0.5,
           xmax = 0.6,
           ymin = 80,
           ymax = 200,
           fill = "yellow",
           alpha = 0.2) +
      annotate("text", 
               label = "Stabbing spear tips",
               size = point_function_group_label_size,
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (200-80) /2 + 80 ) +
  coord_cartesian(xlim = c(-0.1,1.2)) +
  theme_bw(base_size = 6) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()
        ) +
  labs(y = "TCSA")

library(cowplot)
plot_grid(plot_mass_cluster + guides(shape = "none"),
          plot_tcsa_mass_cluster,
          labels = "AUTO",
          label_size = 9,
          rel_widths = c(0.5, 1))

```

## Variation in TCSA values by raw material

@fig-tcsa-raw-materials shows that about half of the stemmed points were made from shale and its TCSA range is widely distributed. The TCSA values of acidic volcanic rocks are commonly skewed lower. Other raw materials including hornfels, rhyolite, and tuff show a wide distribution, similar to shale. The category of "Other" raw materials includes porphyry, trachyte, felsite, chert, quartz, quartzite, granite, mudstone, and unidentified rocks. Overall, there is no clear pattern of TCSA values among different raw materials.

```{r}
#| label: fig-tcsa-raw-materials
#| fig-cap: "TCSA values by lithic raw material."
#| fig-width:   4.4
#| fig-height:  3.2

# Facet by raw materials

 ggplot(tcsa_raw,
              aes(y = min_tcsa, 
                  x = label_raw)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Raw materials", 
       y = "TCSA") +
  theme(axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        text = element_text(size=8)) +
  scale_x_discrete(guide = guide_axis(angle = 45))+
  geom_quasirandom(alpha = 0.4,
                   size = 1.3) +
   annotate("rect", 
           xmin = 6.5,
           xmax = 7,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      annotate("rect", 
           xmin = 6.5,
           xmax = 7,
           ymin = 17,
           ymax = 47,
           fill = "blue",
           alpha = 0.2) +
      annotate("rect", 
           xmin = 6.5,
           xmax = 7,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      annotate("rect", 
           xmin = 6.5,
           xmax = 7,
           ymin = 42,
           ymax = 90,
           fill = "purple",
           alpha = 0.2) +
     annotate("rect", 
           xmin = 6.5,
           xmax = 7,
           ymin = 80,
           ymax = 200,
           fill = "yellow",
           alpha = 0.2) 
```

@fig-tcsa-size shows positive relationships between TCSA, raw material and artifact weight. There is a stronger correlation between TCSA and artifact weight for acidic volcanic points (i.e. points are closer to the regression line), whereas the correlation for shale stemmed points is weaker. The other raw materials show various distribution patterns around the regression line, confirming that raw material appears not to have strongly influenced TCSA (see also @fig-tcsa-raw-materials).

```{r}
#| label: fig-tcsa-size
#| fig-cap: "Artifact size and TSCA values by lithic raw material"
#| fig-height:  3.5

ggplot(tcsa_mass_no_na,
       aes(`Weight(g)`,
           min_tcsa,
           colour = raw_materials_top)) +
  geom_point(size = 1.5) +
    labs(x = "Weight (g)", 
       y = "TCSA",
       colour='Raw material') +
  geom_smooth(method=lm, se=FALSE, size = .7)+
  theme_bw(base_size = 10) +
  theme(
  legend.title = element_text(size = 6.3),
  legend.text = element_text(size = 5.3),
  legend.position = c(.8,.3),
  legend.key.height = unit(4.8, 'mm'),
  legend.key.width = unit(2.5, 'mm'))
```

```{r}
# p-value
# simple version
# p_lgm <- round(summary(aov(min_tcsa ~ as.factor(mis_stage), tcsa_site_ass_mis))[[1]][[5]][1], digits = 2)

# aggregate all the MIS 3 sites, MIS 2 sites
tcsa_site_ass_mis <- tcsa_site_ass %>% 
                      mutate(mis_stage = case_when (
                        age_ka > 26.5 ~ 'Before' , age_ka > 19 ~ 'LGM', age_ka < 19 ~ 'After'))

library("car")
# devtools::install_github('achetverikov/apastats',subdir='apastats')
library("apastats")
p_lgm_test <- Anova(lm(min_tcsa ~ as.factor(mis_stage), tcsa_site_ass_mis))
# describe.Anova(p_lgm_test, 1, 2)

three_time_p_pretty <- apastats::describe.Anova(p_lgm_test, 1, 2)

############ test CV ########### to present
library(raster)
# cv for LGM for example
data_lgm <- tcsa_site_ass_mis %>% 
  filter(mis_stage=='LGM')

cv_lgm <- cv(data_lgm$min_tcsa)

```

## Temporal patterns of TCSA values

@fig-tcsa-radio-carbon shows `r n_sites_with_radiocarbon_ages` assemblages with stemmed points in chronological order (panel A). Assemblages discarded before LGM tend to possess more stemmed points, such as Suyanggae. These assemblages have a wide range of TCSA values, indicating multiple likely uses for stemmed points. Overall, there is no clear pattern in these assemblages over time. The TCSA range varies depending on the assemblage.

In panel B, @fig-tcsa-radio-carbon shows the distribution of TSCA values from assemblages dating to before, during, and after the LGM, in order to explore the impact of climate change on the likely uses of the artifacts. Our results show that stemmed points were made predominantly before LGM and only a few left after LGM. Each category shows wide variation in LGM, which indicates diverse uses of stemmed points. While the median TSCA value for assemblages discarded during the LGM is higher than assemblages from earlier and later, there is no statistically significant difference in TCSA values across the three periods (`r three_time_p_pretty`).

```{r}
#| label: fig-tcsa-radio-carbon
#| fig-cap: "A: Distribution of TCSA values over time. The light gray shade indicates Marine isotope stages (MIS) 3, the gray shade indicates MIS 2 and the darkest shade indicates the duration of LGM. B: Distribution of TCSA values grouped by LGM event"
#| fig-height:  5

temporal_tcsa <- ggplot(tcsa_site_ass, 
       aes(y = min_tcsa,
           x = reorder(label_site, -age_ka))) +
   geom_boxplot(size = 0.2) +
   theme_minimal() +
   theme(legend.position="none") +
   labs(y = "TCSA", 
        x = "") +
  theme(axis.text.x = element_text(size = 5),
        axis.text.y = element_text(size = 5),
        text = element_text(size=8)) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_y_continuous(breaks=seq(50, 200, 50))

# add shady box for MIS stage

all_time <- 
  temporal_tcsa +
 annotate("rect", 
           xmin = 0,
           xmax = 12.2,
           ymin = 240,
           ymax = 260,
           fill = "lightgray",
           alpha = 0.2) +
  annotate("text", 
               size = 3,
               label = "MIS 3",
               hjust = 0,
           x = 5,
           y = 250) +
   annotate("rect", 
           xmin = 12.2,
           xmax = 24.5,
           ymin = 240,
           ymax = 260,
           alpha = 0.2) +
  annotate("text", 
               size = 3,
               label = "MIS 2",
               hjust = 0,
           x = 17,
           y = 250) +
   annotate("rect", 
           xmin = 13.5,
           xmax = 21,
           ymin = 220,
           ymax = 240,
           alpha = 0.4) +
  annotate("text", 
               size = 2.1,
               label = "LGM",
               hjust = 0,
           x = 17.3,
           y = 230)

# box plot for each MIS stage
three_time <-  ggplot(tcsa_site_ass_mis, 
       aes(y = min_tcsa,
           x = factor(mis_stage, level = c('After', 'LGM', 'Before')))) +
   geom_boxplot(size = 0.4) +
   theme_minimal() +
   theme(legend.position="none", 
         axis.text.x = element_text(size = 3.5),
        axis.text.y = element_text(size = 3.5),
        text = element_text(size=5)) +
    labs(y = "TCSA", 
        x = "Time Period") +
  geom_quasirandom(alpha = 0.4,
                   size = 1.1) +
    scale_x_discrete(limits=rev) +
     annotate("rect", 
           xmin = 3.5,
           xmax = 3.8,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      annotate("rect", 
           xmin = 3.5,
           xmax = 3.8,
           ymin = 17,
           ymax = 47,
           fill = "blue",
           alpha = 0.2) +
      annotate("rect", 
           xmin = 3.5,
           xmax = 3.8,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      annotate("rect", 
           xmin = 3.5,
           xmax = 3.8,
           ymin = 42,
           ymax = 90,
           fill = "purple",
           alpha = 0.2) +
     annotate("rect", 
           xmin = 3.5,
           xmax = 3.8,
           ymin = 80,
           ymax = 200,
           fill = "yellow",
           alpha = 0.2)

plot_grid(all_time, 
          three_time, 
          labels = "AUTO",
          label_size = 9,
          ncol = 1,
          rel_heights = c(2.7, 2))

```

## Spatial patterns of TCSA values

We computed artifact TCSA values for `r n_sites` sites that contain more than five stemmed points to observe variation between sites (@fig-tcsa-per-sites). Sites with fewer than five stemmed points were grouped under the category of "Other". Among the sites in our sample, Suyanggae has the most stemmed points and the widest range of TCSA values. This suggests that people made stemmed points for accomplishing a variety of tasks at Suyanggae site. Nosanri, Sibuk, and Wolpyeng have a narrower range of lower values, which indicates less diverse likely uses for stemmed points. Yongsujaeul shows a narrower range but higher TCSA values. Yongsandong has the second highest number of stemmed points and shows two clusters of TCSA values.

```{r}
#| label: fig-tcsa-per-sites
#| fig-cap: "TCSA values by archaeological site."
#| fig-height:  3

 ggplot(tcsa_site,
              aes(y = min_tcsa, 
                  x = label_site)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "", 
       y = "TCSA") +
  theme(axis.text.x = element_text(size = 5)) +
  scale_x_discrete(guide = guide_axis(angle = 45))+
  geom_quasirandom(alpha = 0.4,
                   size = 1.6)+
   annotate("rect", 
           xmin = 7.5,
           xmax = 8,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      annotate("rect", 
           xmin = 7.5,
           xmax = 8,
           ymin = 17,
           ymax = 47,
           fill = "blue",
           alpha = 0.2) +
      annotate("rect", 
           xmin = 7.5,
           xmax = 8,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      annotate("rect", 
           xmin = 7.5,
           xmax = 8,
           ymin = 42,
           ymax = 90,
           fill = "purple",
           alpha = 0.2) +
     annotate("rect", 
           xmin = 7.5,
           xmax = 8,
           ymin = 80,
           ymax = 200,
           fill = "yellow",
           alpha = 0.2)
```

```{r}
######Tukey test######
#vegetation region#

korean_archaeological_site_locations_tcsa <- 
  tcsa %>% 
  left_join(mydata_ages) %>% # add age data
  group_by(Assemblage) %>% 
  mutate(Site_name_to_join = str_squish(Site_name)) %>% 
  left_join(korean_archaeological_site_locations %>% 
              mutate(Site_name_to_join  = str_squish(Site_name)), # deleted invisible space in the data set probably created during data collecting
                     by = "Site_name_to_join") %>% 
  drop_na(min_tcsa) %>% 
  drop_na(age)

library(sf)

############ explore tcsa with MODERN eco_regions #######
library(sf)
veg_all <- read_sf(here("analysis/data/gis_data/eco_re_two.shp"))
veg_all <- st_transform(veg_all, crs = st_crs(4326))
# ggplot() +
#   geom_sf(data = veg_all) +
#   coord_sf()
pts_all <- st_as_sf(korean_archaeological_site_locations_tcsa,
                coords = c("long_dd.y", "lat_dd.y"),
                crs = st_crs(veg_all))
# ggplot() +
#   geom_sf(data = pts_all)
sf_use_s2(FALSE) # to prevent error. Check here for more info: https://github.com/r-spatial/sf/issues/1762
points_in_veg_zones_all <- 
st_join(veg_all, 
        pts_all, 
        join = st_contains)
eco_plot <- ggplot() +
  geom_sf(data = veg_all, aes(fill = "grey")) +
   geom_sf_text(data = veg_all, aes(label = ecoregion), size =1.5)+
  geom_sf(data = pts_all, aes(color = min_tcsa), size = 1.1) +
  theme_minimal() +
  scale_fill_brewer(palette="Greys") +
  labs(color="TCSA") +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, size = 4.5))+
  theme(axis.text.y = element_text(size = 4.5))+
  theme(axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
  theme(legend.position = "none")

eco_box <- ggplot(points_in_veg_zones_all) +
  aes(ecoregion,
      min_tcsa) +
  geom_boxplot(size = 0.2) +
  geom_quasirandom(alpha = 0.3,
                   size = 1.3) +
  xlab("Vegetation") + 
  ylab("TCSA") +
  theme_minimal() +
    theme(axis.text.x = element_text(size = 4.5)) +
  theme(axis.text.y = element_text(size = 4.5)) +
   theme(axis.title.x = element_text(size = 6),
          axis.title.y = element_text(size = 6)) 
# explore artifacts in certain zone, related to Tukey's test
tukey_veg <- points_in_veg_zones_all %>% 
  filter(ecoregion == "ST") %>% 
  dplyr::select(ecoregion, Assemblage, min_tcsa, raw_materials_top, age_ka)

############ explore tcsa with MODERN eco_regions & forest zones #######
veg_forest <- 
  read_sf(here("analysis/data/gis_data/eco_forest_two.shp"))
veg_forest <- 
  st_transform(veg_forest, crs = st_crs(4326))

# ggplot() +
#   geom_sf(data = veg_forest) +
#   coord_sf()
pts_forest <- 
  st_as_sf(korean_archaeological_site_locations_tcsa,
                coords = c("long_dd.y", "lat_dd.y"),
                crs = st_crs(veg_forest))
# ggplot() +
#   geom_sf(data = pts_forest)
#sf_use_s2(FALSE)
points_in_veg_zones_forest <- 
st_join(veg_forest, 
        pts_forest, 
        join = st_contains)
for_plot <- ggplot() +
  geom_sf(data = veg_forest, aes(fill = "grey")) +
  geom_sf_text(data = veg_forest, aes(label = Zone), size =1.5)+
  geom_sf(data = pts_forest, aes(color = min_tcsa), size = 1.1) +
  theme_minimal()+
  scale_fill_brewer(palette="Greys") +
  labs(color="TCSA") +
  guides(fill = "none") +
  theme(axis.text.x = element_text(angle = 45, size = 4.5)) +
  theme(axis.text.y = element_text(size = 4.5)) +
  theme(axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
  theme(legend.title = element_text(size= 6),
        legend.key.height = unit(4.5, 'mm'),
        legend.key.width = unit(2, 'mm'),
        legend.text = element_text(size = 4))

# get rid of zones that has no artifact
points_in_veg_zones_forest_box <- 
  points_in_veg_zones_forest %>% 
  drop_na(Region) 
for_box <- ggplot(points_in_veg_zones_forest_box) +
  aes(Zone,
      min_tcsa) +
  geom_boxplot(size = 0.2) +
  geom_quasirandom(alpha = 0.3,
                   size = 1.3) +
  xlab("Ecoregion") + 
  ylab("TCSA") +
  theme_minimal() +
      theme(axis.text.x = element_text(size = 4.5)) +
  theme(axis.text.y = element_text(size = 4.5)) +
   theme(axis.title.x = element_text(size = 6),
          axis.title.y = element_text(size = 6))

######Tukey test######
#vegetation region#
points_in_veg_zones_all$ecoregion = as.factor(points_in_veg_zones_all$ecoregion)
eco_aov <- aov(min_tcsa ~ ecoregion, points_in_veg_zones_all)
eco_lgm_test <- Anova(lm(min_tcsa ~ ecoregion, points_in_veg_zones_all))
eco_lgm_test_pretty <- apastats::describe.Anova(eco_lgm_test, 1, 2)
eco_tur <- TukeyHSD(eco_aov)

# plot Tukey HSD
eco_tur_box <- 
eco_tur$ecoregion %>% 
  as_tibble(rownames = "zones") %>% 
  ggplot() +
  geom_linerange(aes(xmin = lwr, 
                     xmax = upr,
                     y = zones),
                 size = 0.4) +
  geom_point(aes(diff,
                 zones),
             size = 1) +
  geom_vline(xintercept = 0, colour = "red") +
  ylab("Pairwise comparison\nof TSCA in different\nvegetation zones") +
  xlab("Differences in means") +
  theme_minimal(base_size = 8) +
  theme(axis.text.x = element_text(size = 4.5)) +
  theme(axis.text.y = element_text(size = 4.5)) +
   theme(axis.title.x = element_text(size = 6),
          axis.title.y = element_text(size = 6))
#eco region#
for_aov <- 
  aov(min_tcsa ~ as.factor(Zone),
      points_in_veg_zones_forest_box)
# summary(for_aov)
for_lgm_test <- 
  Anova(lm(min_tcsa ~ as.factor(Zone),
           points_in_veg_zones_forest_box))
for_lgm_test_pretty <- 
  apastats::describe.Anova(for_lgm_test, 1, 2)

for_tur <- TukeyHSD(for_aov)
# plot Tukey HSD
for_tur_box <- 
for_tur$`as.factor(Zone)` %>% 
  as_tibble(rownames = "zones") %>% 
  mutate(zones = fct_reorder(zones, diff)) %>% 
  ggplot() +
  geom_linerange(aes(xmin = lwr, 
                     xmax = upr,
                     y = zones), 
                 size = 0.4) +
  geom_point(aes(diff,
                 zones),
             size = 1) +
  geom_vline(xintercept = 0, colour = "red") +
  ylab("Pairwise comparison\nof TSCA in different\necological zones") +
  xlab("Differences in means") +
  theme_minimal(base_size = 8) +
  theme(axis.text.x = element_text(size = 4.5)) +
  theme(axis.text.y = element_text(size = 4.5)) +
   theme(axis.title.x = element_text(size = 6),
          axis.title.y = element_text(size = 6))

```

In order to explore spatial patterns and the relationship between the role of stemmed points and the environment, we located individual assemblages on maps that depict eco-regional zones. Our results show that stemmed points are only located in certain zones, Central Temperature Zone (CT) of the vegetation map or Central inland (CI) (@fig-gis-tcsa-sites panel A & D) of the eco-region map. We observed that there are statistically significant differences in the distributions of TCSA values between some zones (Vegetation zones: `r eco_lgm_test_pretty`, Ecoregion zones: `r for_lgm_test_pretty`). The Tukey's HSD results show that TCSA values from the South Temperature Zone (ST) zone are significantly different from the other two vegetation zones (@fig-gis-tcsa-sites panel C). Looking into this further, we see that there are only six artifacts from the ST zone, with only three of these having TSCA values above 175. The other three have TCSA values lower than 100, similar to the other zones (@fig-gis-tcsa-sites panel B). Given the small number of artifacts with extreme TCSA values in the ST zone, we hesitate to conclude this is an archaeologically significant pattern of different TCSA values. The Tukey HSD pairwise differences show no pairs of eco-regions with significant differences in TCSA values, likely due to the small number of points in several zones (@fig-gis-tcsa-sites panel F). Overall, our results indicate no clear patterns in the distribution of TCSA values across the vegetation and eco-regional zones.

```{r}
#| label: fig-gis-tcsa-sites
#| fig-cap: "TCSA values by assemblages and ecoregions. A: Modern vegetation map defined by temporature and precipitation. The individual points indicate assemblages. B: TCSA distribution for vegitation zones. The individual points indicate artifacts. C: Tukey test result for vegitation zones. D: Ecoregion map based on geographical boundaries and ecological conditions (modified the map from @lee2008relationship). E: TCSA distribution for ecoregion zones, excluding zones with no artifacts. F: Tukey test result for ecoregion zones"
#| fig-height:  5

plot_grid(eco_plot, 
          for_plot, 
          eco_box, 
          for_box, 
          eco_tur_box, 
          for_tur_box,
          labels = c('A', 'D', 'B', 'E', 'C', 'F'),
          label_size = 9,
          ncol = 2)

```

# Discussion

To investigate the role that stemmed points played in the technological transition during the Korean Late Paleolithic, we computed the TCSA metric to infer likely uses by comparing the results to the TCSA range from other archaeological and ethnographic cases (@tbl-tcsa-ranges). Our main questions were: What were the best-fit ballistic probabilities for the stemmed points if they were hafted as weapon tips? How diverse were their likely uses? What are the temporal and spatial patterns of stemmed point uses?

Our results show that javelin tips and stabbing spear tips are the most probable ballistic uses for stemmed points. There are a few stemmed points in our sample that may have been used as dart tips and un-poisoned arrow tips. None have TCSA values in the range of poison arrow tips (@fig-tcsa-all-sp). @lee2019were computed TCSA for 10 stemmed points from Jingeuneul site, which we were not able to include in this research due to the unavailability of their data, and concluded that their TCSA values fall within the range of North American dart tips and arrowheads. In general, we found a wider range of TCSA in our sample, indicating that stemmed points might play diverse roles in foraging toolkits.

TCSA can be impacted by other factors such as raw materials and portability. @eren2022plains explain that smaller TCSA could be the result of pursuing production economy and transport efficiency. We explored the relationship between TCSA and raw material types and weight in order to examine these factors. We observed no clear pattern for raw materials (@fig-tcsa-raw-materials) but weight (@fig-artifact-size-cluster), though weight is correlated with raw materials (@fig-tcsa-size). We speculate this might be due to the higher availability of raw materials such as shale on hornfels on the landscape. Therefore we could assume that raw materials also associated with TCSA value. We found that different clusters of weight are matched with different types of weapon tips (@fig-artifact-size-cluster).

Our results shows that TCSA values vary between assemblages with few discernible temporal patterns in function (@fig-tcsa-radio-carbon panel A). Stemmed points were used most frequently before the LGM (@fig-tcsa-radio-carbon panel B). While we predicted the greatest variation in TCSA values during the LGM, our results show that variability decreased during the LGM, in comparison to the previous period. This could be explained by the overall decline in stemmed points usage.

We found that stemmed points are most frequently located in a small number of eco-regional zones such as the Central Temperature zone with few clears pattern in TCSA function across the zones (@fig-gis-tcsa-sites). As predicted, TCSA values in inland areas, such as vegetation zone CT, show higher variability. Based on these results, we assume that stemmed points performed a wider range of tasks in low-productivity patches. Both temporal and spatial analyses show that stemmed points were likely more adapted for particular environments, rather than specific climate contexts. @prates2022changes claimed that fishtail projectile points in South America was used for megafauna hunting and contributed to their extinction by proving strong relationship between the temporal density and spatial distribution of megafauna and the projectile points. We similarly found that stemmed points are more densly distributed in certain environments in South Korea, perhaps also relating to the distribution of fauna.

Overall, our temporal and spatial patterns of TCSA show a wide range of TCSA values throughout the Late Paleolithic period and between eco-regional zones. We confirmed the widest TCSA range in a single site, Suyanggae (@fig-tcsa-per-sites). As a consequence, our results are consistent with our second scenario in which stemmed points are described as multifunctional tools, suggesting that people created stemmed points in response to unexpected or varying circumstances in their habitats, similar to how @eren2022plains explain Clovis points.

# Conclusion

Considering the importance of stemmed points for the technological transition during the Late Paleolithic in Korea, our research examines the likely uses for stemmed points by asking three research questions: What were the best-fit ballistic probabilities for the stemmed points if they were hafted as weapon tips? How diverse were their likely uses? What are the temporal and spatial patterns of stemmed point use? We applied the tip cross-sectional area (TCSA) metric because it has been used as a ballistically relevant standard to discriminate different likely use classes of projectile points and it requires only a few measurements on a stone artifact (i.e. width and thickness) to compute the metric. We calculated TCSA for a total of `r min_n_points` stemmed points from `r n_ass` assemblages unearthed from `r n_sites` sites. Then we examined the TCSA values with other variables including raw materials, weight, radiocarbon dates, and site locations. Drawing on evolutionary theory, we premised that the way of using stemmed points is likely to reflect the socio-environmental circumstances that people encountered and managed. We examined the possible impact of LGM on stemmed points and their distribution across eco-regional zones.

According to the different weapon-delivery systems that can be inferred from TCSA values, the majority of stemmed points from the Korean Palaeolithic were probably used as javelin tips and stabbing spear tips. In general, though, we noted a wide range and also differing distributions of TCSA values in each assemblage. Therefore, we conclude that stemmed points had diverse roles during the Late Paleolithic. We conclude that people frequently encountered unfamiliar situations and used stemmed points as a multifunctional tool to carry out multiple tasks. We are aware that discriminating the likely use of small numbers of projectile points could be arbitrary [@erlandson2014darts]. Since TCSA covers the critical elements of projectiles, flight and penetration dynamics (i.e. increased or decrease by shape of tip and cross section), we still consider it a useful metric for discriminating between weapon tip tools [@hughes1998getting; @sitton2020tip; @lombard2021variation]. As part of our future research, we will conduct use-wear analyses of stemmed points and experimental research in an effort to elucidate the detailed function of stemmed points and technological links between hunter-gatherer groups.

# Acknowledgements

<!-- The following line inserts a page break  -->

# CRediT authorship contribution statement

Gayoung Park: Software, Validation, Formal analysis, Resources, Data curation, Writing - original draft, Writing - Review & Editing, Visualization, Project administration.

Marlize Lombard: Conceptualization, Methodology, Supervision

Donghee Chong: Data curation

Ben Marwick: Software, Validation, Formal analysis, Investigation, Writing - Review & Editing, Visualization, Supervision

\newpage

# References {.unnumbered}

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```